A Database Management System for a Geospatially-Enabled Interactive Game

### 2018-12-11 paper/MSDS7330DBMSforGeospatiallyEnabledGame.tex
Small bug fix
Addresses multiple issues in High-level system overview section

\item The database must be able to store the player’s information, including login ID, email, and country of location.
\item The database must be able to represent GPS location data. An important aspect when dealing with geospatial data is how to represent the data efficiently in order to reduce network traffic and storage requirements. A geospatial location is defined by its latitude and longitude coordinates. There are different formats in which these coordinates can be represented. The most common format for many applications (such as Google Maps and Geographical Information Systems or GIS) is the signed degrees format in which latitudes range from -90 to 90, and longitudes from -180 to 180.
\item The database can update the player's location. For the purpose of supporting the gameplay, it is important to update player’s locations when they move to a new destination where collectible objects are available. The database system must ensure the consistency and accuracy of the data even when the information of many closely-located players is been updated frequently. 
\item The location data must be stored with enough precision to support the actions of collecting objects. Latitude and longitude data are more precise when they use more decimals. Using 8 decimal places would give an exact location within 1 millimeter. We can achieve our desired precision using XX decimals. This information is important because it determines, along with the number of players, how much data is stored in the database. If the amount is very large, some sort of data compression would be necessary.
\item The system must have a function to calculate scores for each player, adding the value of the collected objects. The system also records subtotals for each object category and recognizes when thresholds are reached, triggering new story segments.
\item The database must be able to interface with the AWS platform where the data is stored, and the middle layer API.

@@

\subsection{High-level system overview}
In order to be able to serve big number of requests in a timely manner and scale well with increasing number of players we need a system that will not only have a well designed, scalable database, but also other components integrated into a single, well thought architecture. Our system is built from 3 major components:

@@

On the client side we use Apple’s Core Location Framework to collect longitude and latitude of a player and Apple’s ARKit to place objects around player. We calculate bearing and distance between player and a object to place objects around given user. To uniquely identify each player we use Vendor ID.
We want our backend to support large number of requests coming from multiple players. To achieve that we use multiple copies of the identical services capable to scale up or down depending on the level of load. Luckily we do not have to think about scalability properties of our hardware since we run our services in AWS Cloud on EC2 virtual instances.  To spread load across multiple copies of the backend service we use AWS load balancer that distributes client’s requests in a round-robin fashion between EC2 instances. 
We use AWS to host our database. 
Layout of all three components are represented schematically in \autoref{schema}.

>>>>

\item The database must be able to store the player’s information, including login ID, email, and country of location.
\item The database must be able to represent GPS location data. An important aspect when dealing with geospatial data is how to represent the data efficiently in order to reduce network traffic and storage requirements. A geospatial location is defined by its latitude and longitude coordinates. There are different formats in which these coordinates can be represented. The most common format for many applications (such as Google Maps and Geographical Information Systems or GIS) is the signed degrees format in which latitudes range from -90 to 90, and longitudes from -180 to 180.
\item The database can update the player's location. For the purpose of supporting the gameplay, it is important to update player’s locations when they move to a new destination where collectible objects are available. The database system must ensure the consistency and accuracy of the data even when the information of many closely-located players is been updated frequently. 
\item The location data must be stored with enough precision to support the actions of collecting objects. Latitude and longitude data are more precise when they use more decimals. Using 8 decimal places would give an exact location within 1 millimeter. We can achieve our desired precision using 6 decimals. This information is important because it determines, along with the number of players, how much data is stored in the database. If the amount is very large, some sort of data compression would be necessary.
\item The system must have a function to calculate scores for each player, adding the value of the collected objects. The system also records subtotals for each object category and recognizes when thresholds are reached, triggering new story segments.
\item The database must be able to interface with the AWS platform where the data is stored, and the middle layer API.

@@

\subsection{High-level system overview}
In order to be able to serve big number of requests in a timely manner and scale well with increasing number of players we need a system that will not only have a well designed, scalable database, but also other components integrated into a single, well thought architecture. Layout of all components of our system are represented schematically in \autoref{schema}.  Our system is built from 3 major parts:

@@

On the client side we have an iOS application that collects longitude and latitude of a smartphone and places objects around player. We calculate bearing, the horizontal angle between the direction of an object and player and distance between player and an object to place stories around given location.

We want our backend to support large number of requests coming from multiple players. To achieve that we host our middle layer and our database on  Amazon Web Services (AWS).  We can use multiple copies of the identical service to scale our system up or down horizontally depending on the level of load.  To spread load across multiple copies of the service we can use AWS load balancer that distributes client’s requests in a round-robin fashion between instances.
